\sysfull

### 2017-08-03 paper/compiler.tex
more progress, but the text still needs a lot of work.

\sys meets these correctness and efficiency requirements with a four-part
analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis},
(iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.

@@

\textbf{Graph Construction:} starting from a plain C code (see example in Figure~\ref{fig:comp_1}) \sys compiler first builds a control flow graph where (i) each vertices is a \emph{basic block} and (ii) each unidirectional edge is a \emph{branch from one basic block to another} (see example in Figure~\ref{fig:comp_2}). All function calls are inlined in the main function prior to building a graph.

Each vertex has a {\em cost}, which represents the energy usage of the basic block. We used the number of LLVM IRs inside a vertex as the energy use metric. When forming a task, each task should contain a basic block so that the sum of cost of all blocks within each tasks should be roughly the same (viz. O1) and does not exceed given energy budget (viz. C1). For our implementation, the programmer should provide the energy budget in compile time defined in number of LLVM IR instructions that the hardware can execute safely in one energy cycle. This number must be found experimentally.

Each edge holds the information about the {\em live variables} across the edge, and if the edge is within any loop, the {\em loop count}. By this information, the compiler tries to minimize the number of protected variable (viz. O2), because placing boundary at an edge means the live variables across that edge need to be protected. Also the compiler avoids placing boundaries at an edge that is within a loop whose loop count is big (viz. O3).

The compiler also calculates a dominance relation \todo{Define dominance relation}{Kiwan} between vertices. This information is used to ensure task's single-entry property (viz. C2).

\textbf{Loop Analysis:} After constructing the graph and assigning the information on a cost and live variables to each vertex and edge, respectively, the compiler begins with analyzing the program loops. The compiler first identifies all loops and then multiplies the loop count by the cost of a vertices within the loop. The reason is, if the loop is inside a task it will use up energy that is roughly multiplied by the loop count compared to code that is outside of a loop. If the resulting cost of a loop is too large\todo{how do you decide on the acceptable loop cost?}{Kiwan}, the compiler places boundary at the back edge of the loop, and reverts the change to the cost of the vertices inside. It means that if the loop is too large to fit in one task, each iteration should be a separate task execution.

@@

\textbf{Tasks Preparation:} After initial boundary placemen with loops analysis, the compiler places additional boundaries if some tasks are still too large. The \sys compiler starts from the beginning of the main function and traverse through the graph in a depth-first search manner. If any part of the graph is found that is too long to fit in the energy bound, it means a boundary must be placed somewhere within. The subset of graph that needs boundary placement within is called a {\em boundary-candidate}.

@@

\textbf{Ensuring Single Entry:} After placing a boundary, the compiler ensures that the newly made task is single-entried by checking dominance relation. That is, every vertex within a task should be dominated by the entry vertex of the task. If this rule is violated, the compiler places additional boundaries to make sure that every task is single entried.

\subsection{Compiler Transform Pass}
\label{sec:compiler_transform_pass}

>>>>

%\sys meets these correctness and efficiency requirements with a four-part
%analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis},
%(iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.

@@

The input to \sys is a standard C program (as shown in Figure~\ref{fig:comp_1})
and the output of \sys's analysis is the CFG, with some vertices marked as {\em
task boundaries}. 
%
The \sys compiler builds the program's CFG
(as shown in Figure~\ref{fig:comp_2}).  To simplify interprocedural analysis,
our prototype of \sys's analysis inlines all function calls into the main
function before building a graph. 
%
\sys uses a cost-based task boundary placement analysis that
considers per-block energy consumption estimates, inter-task variable liveness,
and the cost of partitioning loops.  We first describe how \sys measures each of these costs 
and then describe \sys's procedure for placing boundaries in the CFG.

\subsubsection{Task Cost Analysis}
\sys's task cost analysis considers estimated task energy cost, variable
liveness, and the impact of splitting loops across tasks.  The result of the
task cost analysis determines where to place boundaries in the CFG.

{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the
CFG a {\em cost}, which is an estimate of the energy consumption of the block.
Precisely modeling the energy cost of arbitrary C code is a difficult problem
and beyond the scope of this work; we used the number of intermediate
representation (IR) instructions in a block as a proxy for energy because
energy cost is likely to tightly correspond  to block length.  Moreover, we
assume as an input to the compiler the maximum number of operations that the
device can execute without exhausting its buffered energy. This parameter is
highly device-dependent, and we assume that a developer or device manufacturer
determines the number empirically, as we did in this work through direct
measurement.  \sys's analysis tries ensure that the estimated energy cost of different tasks
is roughly the same and does not exceed given energy budget.

{\noindent \bf Live variable cost.}
Each CFG edge is associated with a set of {\em live variables} that are
accessible in both of the blocks that define the edge.  Variables that are live
across an edge between two tasks must be protected by \sys's memory
virtualization mechanism and each live, inter-task variable increases the run
time overhead for a boundary placement.  \sys's analysis tries to minimize the
overhead of protected variables by placing boundaries to minimize the number of
inter-task live variables.

{\noindent \bf Loop splitting cost.}
and if the edge is within any loop, the {\em loop count}. 
Also the compiler avoids placing boundaries at an edge that is
within a loop whose loop count is big.

After constructing the graph and assigning the information on a cost and live
variables to each vertex and edge, respectively, the compiler begins with
analyzing the program loops. The compiler first identifies all loops and then
multiplies the loop count by the cost of a vertices within the loop. The reason
is, if the loop is inside a task it will use up energy that is roughly
multiplied by the loop count compared to code that is outside of a loop. If the
resulting cost of a loop is too large\todo{how do you decide on the acceptable
loop cost?}{Kiwan}, the compiler places boundary at the back edge of the loop,
and reverts the change to the cost of the vertices inside. It means that if the
loop is too large to fit in one task, each iteration should be a separate task
execution.

@@

\subsubsection{Task Boundary Placement}
\textbf{Tasks Preparation:} After initial boundary placemen with loops analysis, the compiler places additional boundaries if some tasks are still too large. The \sys compiler starts from the beginning of the main function and traverse through the graph in a depth-first search manner. If any part of the graph is found that is too long to fit in the energy bound, it means a boundary must be placed somewhere within. The subset of graph that needs boundary placement within is called a {\em boundary-candidate}.

@@

\textbf{Ensuring Single Entry:}

The compiler also calculates a dominance relation \todo{Define dominance
relation}{Kiwan} between vertices. This information is used to ensure task's
single-entry property.  After placing a boundary, the compiler ensures that the
newly made task is single-entried by checking dominance relation. That is,
every vertex within a task should be dominated by the entry vertex of the task.
If this rule is violated, the compiler places additional boundaries to make
sure that every task is single entried.

\subsection{Task Transformation}
\label{sec:compiler_transform_pass}
